\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-04-21 doc/sw-system.tex
New text (temporarily in blue) on future surveys, addresses #100 and #139

For example, a $10^4$ square degree survey containing $10^8$
photometrically-selected massive galaxies and $10^5$ lenses could only be
searched by 10 inspectors at a mean rate of 1 galaxy per second and 10
inspections per galaxy in about 14 years. Reducing the inspection time by a
factor of 400 to two weeks would require a robot to reduce the target sample
to 25 per square degree. However, at this point the required purity, 40\%,
would very likely require the average classification time per object to be
more like 10 seconds per object. Hiring 10 inspectors to assess complex images
full time for five months may not be the most cost-effective or
reliable strategy. Alternatively, a team of $10^6$ inspectors could, in
principle, make the required $10^9$ image classifications, $10^3$ each, in a
few weeks; robotically reducing the target list would lead to a proportional
decrease in the required team size.

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been achieved by the online citizen science
project \PH \citep{SchwambEtal2012}.

\PH was designed to enable the discovery of transiting exoplanets in data taken
by the Kepler satellite. A community of over 200,000 inspectors from the general
public found, after each undergoing a small amount of training, over 40 new
exoplanet candidates. They achived this   by visually inspecting the Kepler
lightcurves that were presented in a custom web-based classification interface
\citep{WangEtal2013}.
The older \GZ morphological classification
project \citep{LintottEtal2008} has also enabled the discovery of rare objects,
via its flexible inspection interface and discussion forum
\citep{LintottEtal2009}. Indeed, several of us (AV,CC,CM,EB,PM,TJ,LW) were
active in an informal \GZ gravitational lens search (Verma et al, in
preparation), an experience which led to the present hypothesis that a
systematic online visual lens search could be successful.

@@

If, as expected, the system efficiency could indeed be improved by a factor of
three via real-time classification analysis, we would expect the characteristic
completion time $\tau_0$  in \Eref{eq:speed} to decrease by roughly the same
factor, suggesting that a dataset of $10^4$~images could be searched for lenses
by a crowd of $10^4$ volunteers in about 6 days.

@@

The second way to scale up the number of high contribution classifiers is to
increase the rate at which new volunteers become dedicated volunteers. Based
on feedback from the wider \SW community, this could potentially be achieved
through closer collaboration with the science team. It's also possible that
the dynamic \StageTwo system proposed in the previous section may act as an
incentive to some volunteers to increase their contribution (while gaining
skill in order to ``reach'' \StageTwo). Reducing the rate at which  new
volunteers lose interest could also play a role. Anecdotally, it seems fairly
common for new volunteers to be wary of classifying at all, for fear of
introducing errors. Better explanation of how their early classifications are
analyzed could help assuage these fears: \Sref{sec:results:crowd} shows that
effectively down-weighting new users' classifications (by setting their
agents' initial confusion matrix elements to those of a random classifier)
leads to best performance, a result which should be of some comfort to the
nervous volunteer.

>>>>

For example, a $10^4$ square degree survey containing $10^8$
photometrically-selected massive galaxies (and $10^5$ lenses) could only be
searched by 10 inspectors (at a mean rate of 1 galaxy per second and 3
inspections per galaxy) in about 5 years. Alternatively,  an automated system
could be asked to produce a much purer sample: if it was able to reach a purity
of 10\%, this would leave  $10^6$ lens candidates (100 targets per square
degree) to be visually inspected.  At this point the average visual
classification time per object could well be more like 10 seconds per object,
requiring the same  team of 10 inspectors to work full time for 20 weeks (to
provide 3 classifications  per lens between them) Neither of these may be the
most cost-effective or reliable strategy. Alternatively, a team of $10^6$
inspectors could, in principle, make the required $10^9$ image classifications,
$10^3$ each, in a few weeks; robotically reducing the target list would lead to
a proportional decrease in the required team size.

Systematic detection of rare astronomical objects by such ``crowd-sourced''
visual inspection has recently been demonstrated by the online citizen science
project \PH \citep{SchwambEtal2012}.
\PH was designed to enable the discovery of transiting exoplanets in data taken
by the Kepler satellite. A community of over 200,000 inspectors from the general
public found, after each undergoing a small amount of training, over 40 new
exoplanet candidates. They achieved this   by visually inspecting the Kepler
lightcurves that were presented in a custom web-based classification interface
\citep{WangEtal2013}.
The older \GZ morphological classification
project \citep{LintottEtal2008} has also enabled the discovery of rare objects,
via its flexible inspection interface and discussion forum
\citep{LintottEtal2009}. Indeed, several of us (AV,CC,CM,EB,PM,TJ,LW) were
active in an informal \GZ gravitational lens search (Verma et al, in
preparation), an experience which led to the present hypothesis that a
systematic online visual lens search could be successful.

@@

\new{If, as expected, the system efficiency could indeed be improved by a factor of
three via real-time classification analysis, we would expect the characteristic
completion time $\tau_0$  in \Eref{eq:speed} to decrease by roughly the same
factor, suggesting that a dataset of $10^4$~images could be searched for lenses
by a crowd of $10^4$ volunteers in about 6 days.}

\new{The futuristic lens-finding problem sketched in \Sref{sec:intro}, where $10^5$
lenses are to be found in a $10^4$ square degree wide field imaging survey, is
approximately representative of the challenge facing both the LSST and Euclid
strong lensing science teams \citep{LSSTSciBook,EuclidSciBook}. What role could
citizen scientists at \SW play in lens searching in the next decade? From
\Eref{eq:speed}, and assuming $\tau_0 = 6$ days (as above), we see that for the
\SW crowd to be able to classify $10^8$ images of  photometrically-selected
massive galaxies and groups in approximately 60 days, it would need to contain
$10^7$ volunteers. This is only  a factor of ten larger than the current size of
the Zooniverse userbase. Alternatively, suppose that automated  lens finding
algorithms are able to select as few as $10^6$ targets for visual inspection (a
sample with 10\% purity). Increasing $\tau_0$ by a factor of three to allow for
more inspection time per subject,  we find that \SW could enable a crowd of just
$10^6$ volunteers to assess them in about three  weeks.}

@@

The second way to scale up the number of high contribution classifiers is to
increase the rate at which new volunteers become dedicated volunteers. Based on
feedback from the wider \SW community, this could potentially be achieved
through closer collaboration with the science team. \new{It's also possible that
dynamically assigning subjects on the basis of the volunteer's skill could act
as an incentive to some volunteers to increase their contribution, although any
such ``gamified'' approach would need to be taken thoughtfully.}  Reducing the
rate at which new volunteers lose interest could also play a role. Anecdotally,
it seems fairly common for new volunteers to be wary of classifying at all, for
fear of introducing errors. Better explanation of how their early
classifications are analyzed could help assuage these fears:
\Sref{sec:results:crowd} shows that effectively down-weighting new users'
classifications (by setting their agents' initial confusion matrix elements to
those of a random classifier) leads to best performance, a result which should
be of some comfort to the nervous volunteer.
